---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# neonaop

The goal of neonaop is to make data from the NEON AOP easier to use. 
The National Ecological Observatory Network (NEON) collects hyperspectral 
imagery via its Aerial Observation Platform (AOP) at a 1 meter spatial 
resolution for 426 different wavelengths. 

## Installation

You can install the development version of neonaop via:

```r
#install.packages('devtools')
devtools::install_github('earthlab/neonaop')
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example, message=FALSE}
library(neonaop)
library(raster)
library(viridis)

path_to_file <- system.file('extdata', 'ex.h5', package = 'neonaop')
r <- hs_read(path_to_file, bands = c(1, 50, 100, 400))
r
```

```{r plot}
plot(r, col = cividis(100), axes = FALSE, box = FALSE)
```
