---
title: "Extracting hyperspectral data for plant species in NEON ground plots"
author: "Maxwell B. Joseph"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extract plant spectra}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

NOT_CRAN <- identical(tolower(Sys.getenv('NOT_CRAN')), "true")
knitr::opts_chunk$set(purl = NOT_CRAN, 
                      eval = NOT_CRAN)
```

This vignette shows how the neonaop package can be combined with the 
[neonUtilities](https://github.com/NEONScience/NEON-utilities/) and 
[geoNEON](https://github.com/NEONScience/NEON-geolocation) packages to 
programmatically access woody vegetation 
data from NEON ground plots, and extract spectra from hyperspectral imagery. 

```{r load-libraries, message=FALSE}
library(raster)
library(neonUtilities)
library(geoNEON)
library(neonaop)
library(sp)
library(tidyverse)
```

First, we need to download vegetation data (data acquisition steps
below are adapted from the documentation in the neonUtilities and geoNEON 
packages):

```{r get-veg-data, warning=FALSE, results='hide', message=FALSE}
zipsByProduct(dpID = "DP1.10098.001", 
              site = "TALL", 
              savepath = ".", 
              check.size = FALSE)
stackByTable("filesToStack10098", folder = TRUE)
```

Then, we can read the mapping data and apparent individual data. 
We will restrict our focus to one year's worth of data, to simplify the process
of matching ground plot data in one year to the same year's imagery.

```{r read-map, warning=FALSE, results='hide', message=FALSE}
vegmap <- "filesToStack10098/stackedFiles/vst_mappingandtagging.csv" %>%
  read_csv %>%
  mutate(year = substr(date, 1, 4)) %>%
  filter(year == '2017') %>%
  def.calc.geo.os("vst_mappingandtagging")

vegind <- read_csv("filesToStack10098/stackedFiles/vst_apparentindividual.csv")
```

Merge mapping and individual data, and filter to live plants with coordinates: 

```{r get-precise-locs, warning=FALSE, results='hide', message=FALSE}
veg <- right_join(vegind, vegmap, 
             by = c("individualID", "namedLocation",
                    "domainID", "siteID", "plotID")) %>%
  filter(!is.na(adjEasting), !is.na(adjNorthing), plantStatus == "Live")
```

Using this spatial data, we can acquire tiles of hyperspectral data that 
intersect the plants that were mapped on the ground.

```{r get-hyperspectral-data, results='hide', message=FALSE}
byTileAOP(dpID = "DP3.30006.001", site = "TALL", year = "2017", 
          easting = veg$adjEasting, northing = veg$adjNorthing, 
          check.size = FALSE)
hs_paths <- list.files(path='.', pattern = 'reflectance.h5', 
                       recursive = TRUE, full.names = TRUE)
```

Now, create a SpatialPointsDataFrame of these points. 

```{r make-spdf}
spdf <- SpatialPointsDataFrame(veg[, c('adjEasting', 'adjNorthing')], 
                               data = veg, 
                               proj4string = CRS(hs_proj4string(hs_paths[1])))
```

Now, let's visualize the extents of these hyperspectral images and the locations
of the mapped plants:

```{r plot-locations-and-extents, fig.width=6, fig.height=4}
extents <- lapply(hs_paths, hs_extent)

plot(do.call(raster::merge, extents), bty = 'n', 
     xlab = 'Easting', ylab = 'Northing')
plot(spdf, add = TRUE)
for (e in extents) {
  plot(e, add = TRUE)
}
```

Each mapped plant location is covered by an extent object for a 
hyperspectral image. 
We can use the `neonaop` package to extract hyperspectral data at these 
locations. 

```{r extract-hs-data}
out <- list()
for (i in seq_along(hs_paths)) {
  res <- hs_extract_pts(hs_paths[i], pts = spdf, bands = 1:426)
  first_band <- grep('^band1', names(res), value = TRUE)[1]
  na_vals <- is.na(res[[first_band]])
  out[[i]] <- res[!na_vals, ]
}
```

Now let's create a tibble for easy plotting:

```{r make-tibble}
hs_df <- lapply(out, as.data.frame) %>% 
  bind_rows %>% 
  as_tibble %>%
  select(uid.x, adjEasting, adjNorthing, plantStatus, scientificName, 
         starts_with('band')) %>%
  filter(plantStatus == 'Live') %>%
  distinct
hs_df
```

And gather the hyperspectral columns (converting the data from wide to long 
form).

```{r make-long}
long_df <- hs_df %>%
  gather(band, reflectance, 
         -starts_with('adj'), -plantStatus, -scientificName, -uid.x) %>%
  separate(band, c('index', 'wavelength')) %>%
  mutate(wavelength = parse_number(wavelength)) %>%
  separate(scientificName, into = c('genus', 'species'), sep = ' ', 
           extra = 'drop', remove = FALSE) %>%
  mutate(genusspecies = paste(genus, species)) %>%
  # filter water vapor bands out
  filter(!between(wavelength, 1340, 1445), 
         !between(wavelength, 1790, 1955))
long_df
```

Finally, we can plot the signatures for each species:

```{r plot-hs, fig.width=6, fig.height=3}
long_df %>%
  ggplot(aes(wavelength, reflectance, group = uid.x)) + 
  geom_point(size = .2, alpha = .5) + 
  xlab('Wavelength (nm)') + 
  ylab('Reflectance') + 
  facet_wrap(~genusspecies) + 
  theme_minimal()
```


```{r clean-files, echo = FALSE}
# Finally, clean up the files that were downloaded:
unlink('DP3.30006.001/', recursive = TRUE, force = TRUE)
unlink('filesToStack10098/', recursive = TRUE, force = TRUE)
```
